---
title: "Collecting Individual BICs"
output: html_document
date: "2025-01-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library # Global Variables

```{r library_and_global_var}
library(MASS) 
library(reshape2) 

vb_files_path = "/Users/anjawu/Code/msc-dsaai_thesis_work/vb/"
summary_path = paste0(vb_files_path, 'MCMC_output/summary/')

model_names <- c('pvl_delta', 'orl', 'vpp')
N <- 7

# Create combinations of all individuals without repetition
participant_numbers <- 1:N
subgroups <- unlist(lapply(1:N, function(n) combn(participant_numbers, n, simplify = FALSE)), recursive = FALSE)
```

# Functions

Individual BIC calculator
```{r bic_function}
### BIC(individual | model) = -2 \cdot loglikelihood + p \cdot log(n)
# where p is the number of parameters and n is the number of participants.
bic_indiv_calc_fn <- function(log_lik, n, n_parms) {
  individual = -2*log_lik
  penalty = n_parms*log(n)
  bic_individual = individual + penalty
  return(bic_individual)
}

```


# Collect all individual BIC

Here is a process to amalgamate all BIC scores per individuals in dataframes

```{r collect_bic}
# create dataframe storage for each individual that has columns as model, rows as subgroups, and values as individual BIC score:
individuals_bic_df <- vector("list", N)
for (i in 1:N) {
  individuals_bic_df[[i]] <- data.frame(
    subgroup = character(0),
    matrix(NA, nrow = 0, ncol = length(model_names),
           dimnames = list(NULL, model_names))
  )
}

for (subgroup in subgroups[8:length(subgroups)]) { 
  print(subgroup)
  # collapse numbers of individuals to be able to call the specific set of results
  individuals <- paste(subgroup, collapse = "")
  N_new <- length(subgroup)
  
  for (m in 1:length(model_names)) {
    # create temporary <model> df to access individual values
    temp_model_df <- readRDS(file = paste0(summary_path, "merged_output_from_BEST_igt_subgroup_", individuals,"_igt_", model_names[m], "_sigmainvgam_top15.rds"))
    
    # this summary accesses the mean and standard deviation for each parameter both population and individual parameters
    fit_summary <- rstan::summary(temp_model_df$fit)$summary
    # print(fit_summary)
    # Get a parameters list based on the model examined; exclude the subjID
    parameter_names_list <- colnames(temp_model_df$allIndPars)[2:ncol(temp_model_df$allIndPars)]
    
    # Now we cycle through each individual
    # Calculate BIC scores for each individual in the subgroup
    for (i in 1:N_new) {
      individual_id <- subgroup[i]
      
      individual_log_lik <- fit_summary[paste0('log_lik[', i, ']'), 'mean']
      individual_bic <- bic_indiv_calc_fn(individual_log_lik, 100, length(parameter_names_list))
      
      # Check if row for the subgroup already exists in the individual's dataframe
      if (!(individuals %in% individuals_bic_df[[individual_id]]$subgroup)) {
        # Add a new row with subgroup name and initialize columns
        new_row <- data.frame(
          subgroup = individuals,
          matrix(NA, nrow = 1, ncol = length(model_names),
                 dimnames = list(NULL, model_names))
        )
        individuals_bic_df[[individual_id]] <- rbind(individuals_bic_df[[individual_id]], new_row)
      }
      
      # Assign the BIC score to the appropriate column
      individuals_bic_df[[individual_id]][individuals_bic_df[[individual_id]]$subgroup == individuals, model_names[m]] <- individual_bic
    }
  }
}
```


## Save BIC values/Read back results

```{r save_bic}
# save:
# for (i in 1:N) {
#   saveRDS(individuals_bic_df[[i]], file = paste0(vb_files_path, "individual_bic_scores/individual_", i, "_bic_df.rds"))
# }

# Read back info:
for (i in 1:N) {
  individuals_bic_df[[i]] <- readRDS(paste0(vb_files_path, "individual_bic_scores/individual_", i, "_bic_df.rds"))
}
```


# Boxplot for each individual

```{r boxplots}
plot_path <- paste0(vb_files_path, 'individual_bic_scores/boxplots/')

for (i in 1:N) {
  p <- individuals_bic_df[[i]] |> 
                      melt(id = 'subgroup', variable.name = 'model', value.name = 'bic')|> 
                      ggplot(aes(x=model, y=bic)) + 
                        geom_boxplot() +
                        labs(title=paste0("Distribution of Individual ", i, " BIC Scores by Model"), x="Model Name", y = "Individual BIC Score") +
    scale_x_discrete(labels=c("PVL Delta","ORL","VPP"))
                                                              
  # ggsave(paste0(plot_path, 'boxplot_for_individual', i, '.png'))
  print(p)
}

```

# Conclusions for each individual:

1) Best fit is ORL
2) Best fit is ORL
3) Equivalent fit for PVL Delta and VPP
4) Best fit is PVL Delta
5) Best fit is PVL Delta
6) Best fit is ORL
7) Best fit is ORL, but has a slight overlap with PVL Delta


## Overall from the box plot:
ORL: 1 2 6 7
PVL Delta: 4 5
VPP: 3 (or could be PVL)